Prediction of Protein-Protein Interactions on the Human and Rice Interactome

### 2020-07-30 nicolas/PPIPL3/ReportV1.tex

%% LyX 2.3.5-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twocolumn,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{float}
\usepackage{graphicx}

@@

\usepackage{babel}
\begin{document}
\title{Enhanced Prediction of Protein-Protein Interactions on the Human and
Rice Interactome with Machine Learning}
\author{Nicolás Antonio López Rozo}
\maketitle
\begin{abstract}
Computational tools offer an alternative to identify unmapped, biologically
significant protein-protein interactions (PPI's). A novel approach
by Barabási et al, based on network paths of length three (L3) on
the interactome, outperforms all existing link prediction methods.
L3 can offer therefore a new approach to reveal a smaller subset of
potential interactions for a more affordable experimental validation.

In this study, the previous approach for link prediction is expanded
by using \texttt{XGBoost} module of Machine Learning (ML). For the
feature extraction from the network, the \texttt{Node2Vec} module
was applied, although L3 information is added as a feature for improving
link prediction. Results show that L3 feature is very important for
the ML procedure, although a significant feature importance lies on
the features retrieved from the network structure using \texttt{Node2Vec}.
The mentioned approaches are compared using the human and the rice
interactomes.
\end{abstract}

@@

For achieving the described results, several networks are used and
compared using state-of-the-art methods, as well as the proposed ones
(CN, A3, L3). For this report, the human interactome was used (\emph{HI-II-14.txt}).
A curated version of the mentioned network (\emph{HI-TESTED.txt})
was also used in order to compare the influence of research bias on
the network predictions. A massive experimental assay was carried
on and its results are consolidated and used to build a validation
network (\emph{HI-III.txt}).

@@

Human interactome data and base source code were downloaded from the
repository of the length-3 degree normalized paths methodology\cite{Kovacs2019}:
the dataset \emph{HI-II-14.txt} and \emph{HI-TESTED.txt} are used
for prediction and the dataset \emph{HI-III.txt} is used for validation.
Rice interactome information was downloaded from the STRING database
\cite{Szklarczyk2019}, corresponding to the \emph{Oryza sativa} subspecies.
The downloaded file was \emph{4530.protein.links.detailed.v11.0.txt}.
and contains more than 8 million PPIs from several resources. For
the purpose of this study and based on the previous work, only PPIs
with evidence from curated databases were used (rows where \emph{databases}
column has a value greater than zero), resulting in a network with
NNNN nodes and MMMM edges.

@@

\noindent \centering{}\includegraphics[width=0.8\columnwidth]{\string"/home/ocin/Documents/02 Asignaturas/10 GraphAnalytics/Project1/TCP_extract\string".eps}
\end{figure}

@@

The selected parameters for the model were:
\begin{itemize}
\item \texttt{max\_depth=3}
\item \texttt{colsample\_bytree=0.6}
\item \texttt{eval\_metric='auc'}
\end{itemize}

\subsection{Result Validation}

@@

\begin{figure}[h]
\caption{Methods Comparison for \emph{HI-II-14.txt}}

\noindent \centering{}\includegraphics[width=1\columnwidth]{\string"/home/ocin/Documents/02 Asignaturas/10 GraphAnalytics/Project1/hi-ii-14.txt\string".eps}
\end{figure}

\begin{figure}[h]
\caption{Methods Comparison for \emph{HI-TESTED.txt}}

\noindent \centering{}\includegraphics[width=1\columnwidth]{\string"/home/ocin/Documents/02 Asignaturas/10 GraphAnalytics/Project1/hi-tested.txt\string".eps}
\end{figure}

@@

\bibliographystyle{plain}
\bibliography{\string"/home/ocin/Documents/02 Asignaturas/refs\string"}

>>>>

%% LyX 2.3.5-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[onecolumn,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{float}
\usepackage{graphicx}

@@

\usepackage{babel}
\begin{document}
\title{Prediction of Protein-Protein Interactions on the Human and
Rice Interactome}
\author{Nicol\'as Antonio L\'opez Rozo}
\maketitle
\begin{abstract}
Networks science provides a set of computational tools to identify unmapped protein-protein interactions (PPI's). Barab\'asi et al. recently introduced a metric (based on network paths of length three) which 
proved to be an important feature in predicting links of the interactome network. In other words, paths of length three (L3) enable us to discover hidden link candidates for experimental validation. This paper extends the previous approach by Barab\'asi et al. using \texttt{XGBoost} module of Machine Learning (ML). We train the model using handcrafted features as well as features learned from embeddings. For the feature extraction from the network, \texttt{Node2Vec} is applied. Results show that while L3 feature remains important for predicting links, so are the features retrieved from the network structure using \texttt{Node2Vec}. The mentioned approaches are compared using the human and the rice interactomes.
\end{abstract}

@@

For achieving the described results, several networks are used and
compared using state-of-the-art methods, as well as the proposed ones
(CN, A3, L3). For this report, the human interactome was used (\emph{HI-II-14}).
A curated version of the mentioned network (\emph{HI-TESTED})
was also used in order to compare the influence of research bias on
the network predictions. A massive experimental assay was carried
on and its results are consolidated and used to build a validation
network (\emph{HI-III}).

@@

Human interactome data and base source code were downloaded from the
repository of the length-3 degree normalized paths methodology~\cite{Kovacs2019}:
the dataset \emph{HI-II-14} and \emph{HI-TESTED} are used
for prediction and the dataset \emph{HI-III} is used for validation.
Rice interactome information was downloaded from the STRING database~\cite{Szklarczyk2019}, corresponding to the \emph{Oryza sativa} subspecies.
The downloaded file was \emph{4530.protein.links.detailed.v11.0.txt}.
and contains more than 8 million PPIs from several resources. For
the purpose of this study and based on the previous work, only PPIs
with evidence from curated databases were used (rows where \emph{databases}
column has a value greater than zero), resulting in a network with
NNNN nodes and MMMM edges.

@@

\noindent \centering{}\includegraphics[width=0.8\columnwidth]{TCP_extract.png}
\end{figure}

@@

The selected parameters for the model were \texttt{max\_depth=3}, \texttt{colsample\_bytree=0.6}
and \texttt{eval\_metric=`auc'}.
\subsection{Result Validation}

@@

\begin{figure}[h]
\caption{Methods Comparison for \emph{HI-II-14}}

\noindent \centering{}\includegraphics[width=1\columnwidth]{hi-ii-14.png}
\end{figure}

\begin{figure}[h]
\caption{Methods Comparison for \emph{HI-TESTED}}

\noindent \centering{}\includegraphics[width=1\columnwidth]{hi-tested.png}
\end{figure}

@@

\bibliographystyle{plain}
\bibliography{\string"refs\string"}
